[Result, but poetic]: Does LLM Use Improve Data Science Education?

### 2024-07-24 paper/paper.qmd
finished multibar self-perception visualizations

---
title: "[Result, but poetic]: Does LLM Use Improve Data Science Education?"
subtitle: "Evidence from a Canadian Undergraduate Statistics Course"
author: 
  - Rohan Alexander
  - Luca Carnegie
thanks: "Code and data are available at: https://github.com/lcarnegie/llms-achievement."
date: today
date-format: long
abstract: "Blah Blah Blah Blah"
format: pdf
number-sections: true
bibliography: references.bib
nocite: |
  @*
---

@@

library(tidyverse)
library(knitr)
library(kableExtra)
library(lintr)

@@

@@

```{r fig-writingease-bar}
#| message: false
#| echo: false
#| warning: false
#| tbl-cap: Student Self-Perception of Writing Ease
#| fig-align: center

#Bar plot of writing ease
survey_data |> ggplot(aes(x = writing_is_easy_for_me)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Agreement to the Statement: \"Writing is Easy for Me\"",
       x = "Response",
       y = "Frequency") +
  theme_minimal() +
  theme(plot.title.position = "plot",
        plot.title = element_text(hjust = 0.5))

>>>>

---
title: "[Result, but poetic]: Does LLM Use Improve Data Science Education?"
subtitle: "Evidence from a Canadian Undergraduate Statistics Course"
author: 
  - Rohan Alexander
  - Luca Carnegie
thanks: "Code and data are available at: https://github.com/lcarnegie/llms-achievement."
date: today
date-format: long
abstract: "Blah Blah Blah Blah"
format: pdf
number-sections: true
bibliography: references.bib
nocite: | #references.bib entries at end, regardless of intext citation
  @* 
---

@@

library(tidyverse)
library(knitr)
library(kableExtra)
library(lintr)
library(tidyr)

@@

# Change PEY to 4th year
survey_data <- survey_data |>
    mutate(
      year = case_when(
        year == "PEY" ~ "4th",
        year == "5th" ~ "5th+",
        year == "6th" ~ "5th+",
        TRUE ~ year
      )
    )

@@

```{r fig-response-time-hist}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Distribution of Survey Response Times
#| fig-align: center

#Trim values to histogram

plot_data <- survey_data |>
    filter(mins_complete < 1000)

plot_data |>
    ggplot(aes(x = mins_complete)) +
    geom_histogram(binwidth = 0.1, fill = "skyblue", color = "black") +
    labs(title = "Distribution of Survey Response Times",
         x = "Time (Mins)",
         y = "Frequency") +
    theme_minimal() +
    theme(plot.title.position = "plot",
          plot.title = element_text(hjust = 0.5))

```{r fig-writing-multibar-1}
#| message: false
#| echo: false
#| warning: false
#| tbl-cap: Student Self-Perceptions on Writing, Part 1
#| fig-align: center

# Create a multi-bar plot for writing perceptions grouped by response

#Select first three statements
writing_perceptions <- survey_data |>
  select(
    writing_is_easy_for_me, 
    i_like_to_write, 
    i_believe_it_is_important_to_be_a_good_writer
  )

writing_perceptions_long <- writing_perceptions |>
  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
  mutate(
    Statements = case_when(
      Statements == "writing_is_easy_for_me" ~ "\"Writing is easy for me\"",
      Statements == "i_like_to_write" ~ "\"I like to write\"",
      Statements == "i_believe_it_is_important_to_be_a_good_writer" ~ "\"I believe it is important\n to be a good writer\"",
      TRUE ~ Statements
    ), 
    Response = case_when(
      Response == "This is a lot like me" ~ "A lot like me",
      Response == "This is very different to me" ~ "Very different from me",
      Response == "This somewhat describes me" ~ "Somewhat describes me",
    )
  ) |>
  filter(!is.na(Response))

#Taken from: https://stackoverflow.com/questions/52297978/decrease-overal-legend-size-elements-and-text
addSmallLegend <- function(myPlot, pointSize = 0.5, textSize = 3, spaceLegend = 0.1) {
    myPlot +
        guides(shape = guide_legend(override.aes = list(size = pointSize)),
               color = guide_legend(override.aes = list(size = pointSize))) +
        theme(legend.title = element_text(size = textSize), 
              legend.text  = element_text(size = textSize),
              legend.key.size = unit(spaceLegend, "lines"))
}

# Make the plot
multi_bar <- writing_perceptions_long |>
  ggplot(aes(x = Response, fill = Statements)) +
  geom_bar(position = "dodge", stat = "count") +
  labs(title = "Student Self-Perceptions on Writing",
       x = "Response",
       y = "Frequency") +
  theme_minimal() +
  theme(plot.title.position = "plot",
        plot.title = element_text(hjust = 0.5),
        legend.position= "bottom", 
      )

multi_bar <- addSmallLegend(multi_bar, pointSize = 0.5, textSize = 10, spaceLegend = 0.8)

multi_bar

# TODO: Fix label, Fix Legend title, Fix legibility of Response

```{r fig-writing-multibar-2}
#| message: false
#| echo: false
#| warning: false
#| tbl-cap: Student Self-Perceptions on Writing, Part 2
#| fig-align: center

# Create a multi-bar plot for writing perceptions grouped by response

# Select last three statements
writing_perceptions <- survey_data |>
  select(
    when_i_edit_it_is_easy_for_me_to_catch_my_mistakes, 
    i_feel_confident_sharing_my_writing, 
    i_am_confident_in_my_overall_writing_ability
  )

writing_perceptions_long <- writing_perceptions |>
  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
  mutate(
      Statements = case_when(
      Statements == "when_i_edit_it_is_easy_for_me_to_catch_my_mistakes" ~ "\"When I edit, it is easy to\n catch my own mistakes\"",
      Statements == "i_feel_confident_sharing_my_writing" ~ "\"I feel confident\n sharing my writing\"",
      Statements == "i_am_confident_in_my_overall_writing_ability" ~ "\"I am confident in\n my own writing ability\"",
      TRUE ~ Statements
    ), 
    Response = case_when(
      Response == "This is a lot like me" ~ "A lot like me",
      Response == "This is very different to me" ~ "Very different from me",
      Response == "This somewhat describes me" ~ "Somewhat describes me",
    )
  ) |>
  filter(!is.na(Response))

#Taken from: https://stackoverflow.com/questions/52297978/decrease-overal-legend-size-elements-and-text
addSmallLegend <- function(myPlot, pointSize = 0.5, textSize = 3, spaceLegend = 0.1) {
    myPlot +
        guides(shape = guide_legend(override.aes = list(size = pointSize)),
               color = guide_legend(override.aes = list(size = pointSize))) +
        theme(legend.title = element_text(size = textSize), 
              legend.text  = element_text(size = textSize),
              legend.key.size = unit(spaceLegend, "lines"))
}

# Make the plot
multi_bar <- writing_perceptions_long |>
  ggplot(aes(x = Response, fill = Statements)) +
  geom_bar(position = "dodge", stat = "count") +
  labs(title = "Student Self-Perceptions on Writing",
       x = "Response",
       y = "Frequency") +
  theme_minimal() +
  theme(plot.title.position = "plot",
        plot.title = element_text(hjust = 0.5),
        legend.position= "bottom", 
      )

multi_bar <- addSmallLegend(multi_bar, pointSize = 0.5, textSize = 10, spaceLegend = 0.8)

multi_bar
```

```{r fig-coding-multibar-1}
#| message: false
#| echo: false
#| warning: false
#| tbl-cap: Student Self-Perceptions on Coding, Part 1
#| fig-align: center

# Create a multi-bar plot for writing perceptions grouped by response

#Select first three statements
writing_perceptions <- survey_data |>
  select(
    coding_is_easy_for_me,
    i_like_to_code,
    i_believe_it_is_important_to_be_a_good_coder
  )

writing_perceptions_long <- writing_perceptions |>
  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
  mutate(
    Statements = case_when(
      Statements == "coding_is_easy_for_me" ~ "\"Coding is easy for me\"",
      Statements == "i_like_to_code" ~ "\"I like to code\"",
      Statements == "i_believe_it_is_important_to_be_a_good_coder" ~ "\"I believe it is important\n to be a good coder\"",
      TRUE ~ Statements
    ), 
    Response = case_when(
      Response == "This is a lot like me" ~ "A lot like me",
      Response == "This is very different to me" ~ "Very different from me",
      Response == "This somewhat describes me" ~ "Somewhat describes me",
    )
  ) |>
  filter(!is.na(Response))

#Taken from: https://stackoverflow.com/questions/52297978/decrease-overal-legend-size-elements-and-text
addSmallLegend <- function(myPlot, pointSize = 0.5, textSize = 3, spaceLegend = 0.1) {
    myPlot +
        guides(shape = guide_legend(override.aes = list(size = pointSize)),
               color = guide_legend(override.aes = list(size = pointSize))) +
        theme(legend.title = element_text(size = textSize), 
              legend.text  = element_text(size = textSize),
              legend.key.size = unit(spaceLegend, "lines"))
}

# Make the plot
multi_bar <- writing_perceptions_long |>
  ggplot(aes(x = Response, fill = Statements)) +
  geom_bar(position = "dodge", stat = "count") +
  labs(title = "Student Self-Perceptions of Coding",
       x = "Response",
       y = "Frequency") +
  theme_minimal() +
  theme(plot.title.position = "plot",
        plot.title = element_text(hjust = 0.5),
        legend.position= "bottom", 
      )

multi_bar <- addSmallLegend(multi_bar, pointSize = 0.5, textSize = 10, spaceLegend = 0.8)

multi_bar

```{r fig-coding-multibar-2}
#| message: false
#| echo: false
#| warning: false
#| tbl-cap: Student Self-Perceptions on Coding, Part 2
#| fig-align: center

# Create a multi-bar plot for writing perceptions grouped by response

#Select first three statements
writing_perceptions <- survey_data |>
  select(
    when_i_check_my_code_it_is_easy_for_me_to_catch_my_mistakes, 
    i_feel_confident_sharing_my_code,
    i_am_confident_in_my_overall_coding_ability
  )

writing_perceptions_long <- writing_perceptions |>
  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
  mutate(
    Statements = case_when(
      Statements == "when_i_check_my_code_it_is_easy_for_me_to_catch_my_mistakes" ~ "\"When I check my code, it is\n easy for me to catch my mistakes\"",
      Statements == "i_feel_confident_sharing_my_code" ~ "\"I feel confident sharing my code\"",
      Statements == "i_am_confident_in_my_overall_coding_ability" ~ "\"I am confident in\n my overall coding ability\"",
      TRUE ~ Statements
    ), 
    Response = case_when(
      Response == "This is a lot like me" ~ "A lot like me",
      Response == "This is very different to me" ~ "Very different from me",
      Response == "This somewhat describes me" ~ "Somewhat describes me",
    )
  ) |>
  filter(!is.na(Response))

#Taken from: https://stackoverflow.com/questions/52297978/decrease-overal-legend-size-elements-and-text
addSmallLegend <- function(myPlot, pointSize = 0.5, textSize = 3, spaceLegend = 0.1) {
    myPlot +
        guides(shape = guide_legend(override.aes = list(size = pointSize)),
               color = guide_legend(override.aes = list(size = pointSize))) +
        theme(legend.title = element_text(size = textSize), 
              legend.text  = element_text(size = textSize),
              legend.key.size = unit(spaceLegend, "lines"))
}

# Make the plot
multi_bar <- writing_perceptions_long |>
  ggplot(aes(x = Response, fill = Statements)) +
  geom_bar(position = "dodge", stat = "count") +
  labs(title = "Student Self-Perceptions of Coding",
       x = "Response",
       y = "Frequency") +
  theme_minimal() +
  theme(plot.title.position = "plot",
        plot.title = element_text(hjust = 0.5),
        legend.position= "bottom", 
      )

multi_bar <- addSmallLegend(multi_bar, pointSize = 0.4, textSize = 8.5, spaceLegend = 0.8)

multi_bar
